---
title: "KAGGLE_FINAL"
output: html_notebook
---

### KNN MODELING
1. Standardizing data for KNN & setting our target variable as categorical
```{r}
Train <- read.csv("FlightTrainNew1.csv")
Test <- read.csv("FlightTestNoYNew1.csv")
Train$Cancelled <- as.factor(Train$Cancelled)

data_OG <- Train

# Removing rows with more than 50% NAs
NA_cols = c()
for(i in 1:ncol(data_OG)) {
  if(sum(is.na(data_OG[[i]])) > (.5 * nrow(data_OG))) NA_cols = c(NA_cols, i)
}

length(NA_cols)
data_NAfree = data_OG[,-NA_cols]

Train <- data_NAfree

# Standardizing data
Train$SCHEDULED_DEPARTURE <- scale(Train$SCHEDULED_DEPARTURE)
Train$SCHEDULED_ARRIVAL <- scale(Train$SCHEDULED_ARRIVAL)
Train$SCHEDULED_TIME <- scale(Train$SCHEDULED_TIME)

Train$Org_airport_lat <- scale(Train$Org_airport_lat)
Train$Org_airport_long <- scale(Train$Org_airport_long)
Train$Dest_airport_lat <- scale(Train$Dest_airport_lat)
Train$Dest_airport_long <- scale(Train$Dest_airport_long)

Test$SCHEDULED_DEPARTURE <- scale(Test$SCHEDULED_DEPARTURE)
Test$SCHEDULED_ARRIVAL <- scale(Test$SCHEDULED_ARRIVAL)
Test$SCHEDULED_TIME <- scale(Test$SCHEDULED_TIME)

Test$Org_airport_lat <- scale(Test$Org_airport_lat)
Test$Org_airport_long <- scale(Test$Org_airport_long)
Test$Dest_airport_lat <- scale(Test$Dest_airport_lat)
Test$Dest_airport_long <- scale(Test$Dest_airport_long)
```

2. knn model using Org_airport_lat, Org_airport_long, Dest_airport_lat, Dest_airport_long, SCHEDULED_DEPARTURE, SCHEDULED_TIME, SCHEDULED_ARRIVAL, MONTH, DAY. k=1.
model using Org_airport_lat, Org_airport_long, Dest_airport_lat, Dest_airport_long, SCHEDULED_DEPARTURE, SCHEDULED_TIME, MONTH gave us the same accuracy rate (100%)
```{r}
library(class)
# Gives us training data accuracy
model.train <- knn(Train[,c(14:19, 24:26)], Train[,c(14:19, 24:26)], cl=Train[,28],k=1)
a1 <- table(Train[,28],model.train)
a1

model.train <- knn(Train[,c(14:19, 24:26)], Test[,c(14:19, 24:26)], cl=Train[,28],k=1)
```

```{r}
library(caret)
library(e1071)
trctrl <- trainControl(method="repeatedcv", number=5)
knn_fit <- train(Cancelled ~ Org_airport_lat + Org_airport_long +
                   Dest_airport_lat + Dest_airport_long +
                   SCHEDULED_DEPARTURE + SCHEDULED_TIME + SCHEDULED_ARRIVAL +
                   MONTH + DAY, data=Train, method="knn",
                 tuneGrid   = expand.grid(k = 1),
                 trControl=trctrl, metric="Accuracy")

```
```{r}
tt <- data.frame(knn_fit[4])
plot(tt[,1], tt[,2], type="b", col="blue", xlab="k Nearest Neighbor", ylab="Accuracy")
knn_fit

```

3. Changes based on looking through the data (Linus' five variables and DIVERTED)
```{r}
# First, check if air_system_delay leads to cancellations
table(Train$Cancelled[Train$AIR_SYSTEM_DELAY != 0])
# When air_system_delay != 0, flight NOT CANCELLED

# Next, check security_delay
table(Train$Cancelled[Train$SECURITY_DELAY != 0])
# When security_delay != 0, flight NOT CANCELLED

# Check for late_aircraft_delay
table(Train$LATE_AIRCRAFT_DELAY)
table(Train$Cancelled[Train$LATE_AIRCRAFT_DELAY != 0])
# When late_aircraft_delay != 0, flight NOT CANCELLED

# Check for weather_delay
table(Train$Cancelled[Train$WEATHER_DELAY != 0])
# When weather_delay != 0, flight NOT CANCELLED

# Check for airline_delay
table(Train$Cancelled[Train$AIRLINE_DELAY != 0])
# When airline_delay != 0, flight NOT CANCELLED

# Make these changes to Chaeyoung's submission
ASD_holder = which(Test$AIR_SYSTEM_DELAY != 0)
SD_holder = which(Test$SECURITY_DELAY != 0)
LAD_holder = which(Test$LATE_AIRCRAFT_DELAY != 0)
WD_holder = which(Test$WEATHER_DELAY != 0)
AD_holder = which(Test$AIRLINE_DELAY != 0)
DI_holder = which(Test$DIVERTED != 0)

# Check number of fixes
length(unique(c(ASD_holder, SD_holder, LAD_holder, WD_holder, AD_holder, DI_holder)))
# 4476 possible fixes

all_fixes = unique(c(ASD_holder, SD_holder, LAD_holder, WD_holder, AD_holder, DI_holder))
```

4. Outputting to a csv file
```{r}
Ob <- c(1:29668)
Cancelled <- as.character(model.1)
sol <- cbind(Ob, Cancelled)
table(Cancelled)

write.csv(sol, "sol.csv", row.names = FALSE)
best = read.csv("sol.csv")
# Check how many possible fixes there are
table(best$Cancelled[all_fixes])
# There are 184 possible "NO"s
best$Cancelled[all_fixes] = "NO"

# Check distribution
table(best$Cancelled)

write.csv(best, "sol2.csv", row.names = FALSE)
```

### RANDOMFOREST
```{r}
library(randomForest)
set.seed(1)

bag=randomForest(Cancelled~Org_airport_lat + Org_airport_long + Dest_airport_lat + Dest_airport_long + SCHEDULED_DEPARTURE + SCHEDULED_TIME + SCHEDULED_ARRIVAL + MONTH + DAY,data=Train,mtry=19,importance=TRUE)
summary(bag)
plot(bag)
print(bag)
```

```{R}
model.1 <- predict(bag, newdata=Test)
Ob <- c(1:29668)
Cancelled <- as.character(model.1)
sol <- cbind(Ob, Cancelled)
table(Cancelled)

write.csv(sol, "sol.csv", row.names = FALSE)

# best = read.csv("sol.csv")
# # Check how many possible fixes there are
# table(best$Cancelled[all_fixes])
# # There are 184 possible "NO"s
# best$Cancelled[all_fixes] = "NO"
# 
# # Check distribution
# table(best$Cancelled)
# 
# write.csv(best, "sol2.csv", row.names = FALSE)
```

```{r}
library(gbm)
set.seed(1)
Train$Cancelled <- ifelse(Train$Cancelled=="YES", 1,0)
boost <- gbm(Cancelled~Org_airport_lat + Org_airport_long + Dest_airport_lat + Dest_airport_long + SCHEDULED_DEPARTURE + SCHEDULED_TIME + SCHEDULED_ARRIVAL + MONTH + DAY,data=Train,distribution="bernoulli")
p.boost <- predict(boost, newdata=Train, type="response")
p.boost1 <- ifelse(p.boost<0.5,0,1)
table(Train$Cancelled, p.boost1)
```
